---
title: "edld610_lab2"
author: "Cameron Kay"
date: "1/13/2019"
output:
  html_document:
    highlight: espresso
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
---

## Write the code to load the data in directly from the web

```{r setup, message = FALSE}
# load packages
library(tidyverse)
library(magrittr)
library(viridis)
library(Hmisc)

# import data
df <- rio::import(file     = "https://github.com/fivethirtyeight/data/raw/master/alcohol-consumption/drinks.csv",
                  setclass = "tibble") %>%
  janitor::clean_names()
```

## Explore the overall distribution of total alcohol consumed

### Histogram

#### Play with the number of bins

```{r histogram}
# histogram with 5 bins
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_histogram(bins = 5, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# histogram with 10 bins
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_histogram(bins = 10, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# histogram with 13 bins
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_histogram(bins = 13, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# histogram with 15 bins
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_histogram(bins = 15, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# histogram with 25 bins
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_histogram(bins = 25, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# histogram with 50 bins
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_histogram(bins = 50, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# histogram with 100 bins
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_histogram(bins = 100, fill = "darkorchid4", color = "gray17") +
  theme_bw()
```

#### Justify the number of bins

I think 13 bins is the correct number of bins for this dataset. If the idea is to give an impression of the underlying population distribution, 13 bins shows the nuance of the data, while not including as much presumably random variation as the 15 bin histrogram. On the other side, the 10 bin histogram seems to hide too much of the underlying distribution. 

### Density Plot

```{r density_plots}
# density plot with a bandwidth of .1
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_density(bw = .1, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# density plot with a bandwidth of .5
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_density(bw = .5, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# density plot with a bandwidth of .9
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_density(bw = .9, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# density plot with a bandwidth of 1
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_density(bw = 1, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# density plot with a bandwidth of 1.5
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_density(bw = 1.5, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# density plot with a bandwidth of 2
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_density(bw = 2, fill = "darkorchid4", color = "gray17") +
  theme_bw()

# density plot with the default bandwidth
ggplot(df, aes(x = total_litres_of_pure_alcohol)) +
  geom_density(fill = "darkorchid4", color = "gray17") +
  theme_bw()

```

#### Justify the bandwidth

Using the same criteria as the histogram, I would go with a bandwidth of .9. I believe the .5 distribution provides too much nuance in the data and may be including too much random variation. A bandwidth of 1 and the default bandwidth, on the other hand, appear to minimize the valleys between the peaks and may hide the trimodal nature of the distribution.

## Reorganize the data so I have a type and servings column.

```{r reorganize_data}
df %<>%
  gather(key = "type", value = "servings", contains("_servings")) %>%
  mutate(type = gsub("_servings", "", type))
```

### Evaluate the distribution of servings by type using at least three different methods

```{r visualize_servings_by_type}
# boxplot
ggplot(df, aes(x = type, y = servings, fill = type)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE) +
  theme_bw()

# violin plot
ggplot(df, aes(x = type, y = servings, fill = type)) +
  geom_violin() +
  scale_fill_viridis(discrete = TRUE) +
  theme_bw()

# dotplot
ggplot(df, aes(x = type, y = servings, fill = type)) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 5, alpha = .7) +
  scale_fill_viridis(discrete = TRUE) +
  theme_bw()
```

### Conclusion

Qualitatively, the average number of servings consumed appears to be lowest for wine and highest for beer. Spirits fall in the middle. Consumption of all types also appears to be positively skewed, with the majority of people consuming low levels of alcohol and a tail of increasingly heavy drinkers. Wine has a particularly skinny tail, with the vast majority of people cosuming little to no wine and some individuals consuming quite a bit of wine. The greatest number of servings for any alcohol type was for spirits.

### Reproduce the plot

```{r reproduce_plot}
# reproduced plot
df %>%
  spread(key = type, value = servings) %>%
  arrange(desc(total_litres_of_pure_alcohol)) %>%
  dplyr::slice(1:25) %>%
  gather(key = "type", value = "servings", beer:wine) %>%
  ggplot(aes(x = fct_reorder(country, total_litres_of_pure_alcohol), 
             y = total_litres_of_pure_alcohol)) + # shouldn't servings be on the y-axis?
    geom_col(fill = "steelblue2", alpha = .8) +
    coord_flip() + 
    facet_grid(. ~ type, labeller = labeller(type = capitalize)) + 
    theme_minimal() +
    labs(title = "Countries with the highest alcohol consumption",
         subtitle = "Top 25 countries displayed",
         y = "Total number of alcohol servings",
         x = "Country")

# reproduced plot with number of servings for each alcohol type on the y-axis
df %>%
  spread(key = type, value = servings) %>%
  arrange(desc(total_litres_of_pure_alcohol)) %>%
  dplyr::slice(1:25) %>%
  gather(key = "type", value = "servings", beer:wine) %>%
  ggplot(aes(x = fct_reorder(country, total_litres_of_pure_alcohol), 
             y = servings)) + 
    geom_col(fill = "steelblue2", alpha = .8) +
    coord_flip() + 
    facet_grid(. ~ type, labeller = labeller(type = capitalize)) +
    theme_minimal() +
    labs(title = "Countries with the highest alcohol consumption",
         subtitle = "Top 25 countries displayed",
         y = "Total number of alcohol servings",
         x = "Country")
```
